---
title: "H1"
author: "Nemish"
date: "April 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r library_import, message=FALSE}
library(ggplot2)
library(stringr)
library(dplyr)
library(corrplot)
```

```{r dataset import}
train = read.csv('train.csv', stringsAsFactors = FALSE)
test = read.csv('test.csv', stringsAsFactors = FALSE)
```

```{r}
df.combined = rbind(within(train, rm('SalePrice')), test)
rownames(df.combined) <- df.combined$Id
df.combined$Id <- NULL
```
count of null values present in the columns of `df.combined`
```{r}
na.cols = which(colSums(is.na(df.combined)) > 0)
sort(colSums(sapply(df.combined[na.cols], is.na)), decreasing = TRUE)
paste('No. of columns with missing values:', length(na.cols) )
```

###helper function for plotting categoric data for easier data visualization
```{r}
plot.categoric = function(cols, df){
  for (col in cols) {
    order.cols = names(sort(table(df[,col]), decreasing = TRUE))
  
    num.plot = qplot(df[,col]) +
      geom_bar(fill = '#00A4CC') +
      geom_text(aes(label = ..count..), stat='count', vjust=-0.5) +
      theme_minimal() +
      scale_y_continuous(limits = c(0,max(table(df[,col]))*1.1)) +
      scale_x_discrete(limits = order.cols) +
      xlab(col) +
      theme(axis.text.x = element_text(angle = 30, size=12))
  
    print(num.plot)
  }
}
```

###Pool chunk
```{r}
pool.cols = names(df.combined)[sapply(names(df.combined), function(x) str_detect(x, 'Pool'))]
df.combined[is.na(df.combined$PoolQC) & (df.combined$PoolArea > 0), c('PoolQC','PoolArea')]

df.combined[,c(pool.cols)] %>%
  group_by(PoolQC) %>%
    summarise(mean = mean(PoolArea), counts = n())

df.combined[2421,'PoolQC'] = 'Ex'
df.combined[2504,'PoolQC'] = 'Ex'
df.combined[2600,'PoolQC'] = 'Fa'
df.combined$PoolQC[is.na(df.combined$PoolQC)] = 'None'

```

###Garage Chunk
```{r}
length(which(df.combined$GarageYrBlt == df.combined$YearBuilt))

idx = which(is.na(df.combined$GarageYrBlt))
df.combined[idx, 'GarageYrBlt'] <- df.combined[idx, 'YearBuilt']

garage.cols = names(df.combined)[sapply(names(df.combined), function(x) str_detect(x, 'Garage'))]
sort(colSums(sapply(df.combined[garage.cols], is.na)), decreasing = TRUE)

df.combined[is.na(df.combined$GarageCond),garage.cols]

df.combined[,c(garage.cols)] %>%
  group_by(GarageType) %>%
    summarise(median0 = median(GarageCars, na.rm=TRUE), median1 = median(GarageArea, na.rm=TRUE), counts = n())

###Treating NA's in Cars and Area field
### for Detchd
df.combined[is.na(df.combined$GarageArea),garage.cols]
df.combined[2577, 'GarageArea'] <- 399.5
df.combined[2577, 'GarageCars'] <- 2

idx = which(((df.combined$GarageArea < 370) & (df.combined$GarageArea > 350)) & (df.combined$GarageCars == 1))
names(sapply(df.combined[idx, garage.cols], function(x) sort(table(x), decreasing=TRUE)[1]))

df.combined[2127,'GarageQual'] = 'TA'
df.combined[2127, 'GarageFinish'] = 'Unf'
df.combined[2127, 'GarageCond'] = 'TA'

idx = which(((df.combined$GarageArea < 410) & (df.combined$GarageArea > 380)) & (df.combined$GarageCars == 2))
names(sapply(df.combined[idx, garage.cols], function(x) sort(table(x), decreasing=TRUE)[1]))
df.combined[2577,'GarageQual'] = 'TA'
df.combined[2577, 'GarageFinish'] = 'Fin'
df.combined[2577, 'GarageCond'] = 'TA'

for (col in garage.cols){
  if (sapply(df.combined[col], is.numeric) == TRUE){
    df.combined[is.na(df.combined[col]), col] = 0
  }
  else{
    df.combined[is.na(df.combined[col]), col] = 'None'
  }
}
```

###Kitchen and Electrical Chunk
```{r}
plot.categoric('KitchenQual', df.combined)
df.combined$KitchenQual[is.na(df.combined$KitchenQual)] = 'TA'
plot.categoric('Electrical', df.combined)
df.combined$Electrical[is.na(df.combined$Electrical)] = 'SBrkr'
```

###Basement Chunk
```{r}
bsmt.cols = names(df.combined)[sapply(names(df.combined), function(x) str_detect(x, 'Bsmt'))]

sort(colSums(sapply(df.combined[bsmt.cols], is.na)), decreasing = TRUE)
df.combined[is.na(df.combined$BsmtExposure),bsmt.cols]
df.combined[c(949, 1488, 2349), 'BsmtExposure'] = 'No'

for (col in bsmt.cols){
  if (sapply(df.combined[col], is.numeric) == TRUE){
    df.combined[sapply(df.combined[col], is.na),col] = 0
  }
  else{
    df.combined[sapply(df.combined[col],is.na),col] = 'None'
  }
}
```

###Exterior Chunk
```{r}
idx = which(is.na(df.combined$Exterior1st) | is.na(df.combined$Exterior2nd))
df.combined[idx,c('Exterior1st', 'Exterior2nd')]
df.combined$Exterior1st[is.na(df.combined$Exterior1st)] = 'Other'
df.combined$Exterior2nd[is.na(df.combined$Exterior2nd)] = 'Other'
```

###SaleType Chunk
```{r}
idx = which(is.na(df.combined$SaleType) )
df.combined[idx, 'SaleCondition']
table(df.combined$SaleCondition, df.combined$SaleType)
df.combined[idx, 'SaleType'] <- 'WD'
```

###Functional Chunk
```{r}
table(df.combined$Functional)
idx = which(is.na(df.combined$Functional))
df.combined[idx, 'Functional'] <- 'Typ'
```

###Utilities Chunk
```{r}
table(df.combined$Utilities)
df.combined$Utilities <- NULL
```

###MS Chunk
```{r}
MS.cols = names(df.combined)[sapply(names(df.combined), function(x) str_detect(x, 'MS'))]

idx = which(is.na(df.combined$MSZoning))
df.combined[idx, 'MSSubClass']
table(df.combined$MSSubClass, df.combined$MSZoning)
df.combined[idx,c('MSZoning','MSSubClass')]
df.combined$MSZoning[c(1916, 2251)] <- 'RM'
df.combined$MSZoning[c(2217, 2905)] <- 'RL'
```

###Msn Chunk
```{r}
df.combined[(is.na(df.combined$MasVnrType)) | (is.na(df.combined$MasVnrArea)), c('MasVnrType', 'MasVnrArea')]

df.combined[,c('MasVnrType','MasVnrArea')] %>%
  group_by(MasVnrType) %>%
    summarise(median0 = median(MasVnrArea, na.rm=TRUE), counts = n())

df.combined[2611, 'MasVnrType'] <- 'BrkFace'

df.combined$MasVnrType[is.na(df.combined$MasVnrType)] = 'None'
df.combined$MasVnrArea[is.na(df.combined$MasVnrArea)] = 0
```

###LotFrontage Chunk
```{r}
lot.by.nbrh <- df.combined[,c('Neighborhood','LotFrontage')] %>%
  group_by(Neighborhood) %>%
  summarise(median = median(LotFrontage, na.rm = TRUE))

idx = which(is.na(df.combined$LotFrontage))

for (i in idx){
  lot.median = lot.by.nbrh[lot.by.nbrh == df.combined$Neighborhood[i],'median']
  df.combined[i,'LotFrontage'] = lot.median[[1]]
}
```

###Fence Chunk
```{r}
table(df.combined$Fence)
df.combined$Fence[is.na(df.combined$Fence)] = 'None'
```

###MiscFeature Chunk
```{r}
df.combined$MiscFeature[is.na(df.combined$MiscFeature)] = 'None'
```

###FirePlace Chunk
```{r}
Fire.cols = names(df.combined)[sapply(names(df.combined), function(x) str_detect(x, 'Fire'))]

which((df.combined$Fireplaces > 0) & (is.na(df.combined$FireplaceQu)))
df.combined$FireplaceQu[is.na(df.combined$FireplaceQu)] = 'None'
```

###Alley Chunk
```{r}
df.combined$Alley[is.na(df.combined$Alley)] = 'None'
```

```{r}
paste('There are', sum(sapply(df.combined, is.na)), 'missing values left')
```

##Numeric or Categorical Feature
```{r}
df.numeric = df.combined[num_features]

num_features = names(which(sapply(df.combined, is.numeric)))
cat_features = names(which(sapply(df.combined, is.character)))

group.df = df.combined[1:1460,]
group.df$SalePrice = train$SalePrice

# function that groups a column by its features and returns the mdedian saleprice for each unique feature. 
group.prices = function(col) {
  group.table = group.df[,c(col, 'SalePrice', 'OverallQual')] %>%
    group_by_(col) %>%
    summarise(mean.Quality = round(mean(OverallQual),2),
      mean.Price = mean(SalePrice), n = n()) %>%
    arrange(mean.Quality, mean.Price)
    
  print(qplot(x=reorder(group.table[[col]], -group.table[['mean.Price']]), y=group.table[['mean.Price']]) +
    geom_bar(stat='identity', fill='#00A4CC') +
    theme_minimal() +
    #scale_y_continuous(labels = dollar) +
    labs(x=col, y='Mean SalePrice') +
    theme(axis.text.x = element_text(angle = 45))) +
    #geom_text(aes(label = sort(median.Price)), vjust = -0.5)
  
  return(data.frame(group.table))
}

# function that maps a categoric value to its corresponding numeric value and returns that column to the data frame
map.fcn = function(cols, map.list, df){
  for (col in cols){
    df[col] = as.numeric(map.list[df.combined[,col]])
  }
  return(df)
}

qual.list = c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)
qual.cols = c('ExterQual', 'ExterCond', 'GarageQual', 'GarageCond', 'FireplaceQu', 'KitchenQual', 'HeatingQC', 'BsmtQual')

df.numeric = map.fcn(qual.cols, qual.list, df.numeric)
```

###BsmtExposure, BsmtFinType1
```{r}
group.prices('BsmtExposure')
bsmt.levels = c('None' = 0, 'No' = 1, 'Mn' = 2, 'Av' = 3, 'Gd' = 4)
df.numeric = map.fcn(c('BsmtExposure'), bsmt.levels, df.numeric)

group.prices('BsmtFinType1')

df.combined[, c('BsmtFinType1', 'BsmtFinSF1')] %>%
  group_by(BsmtFinType1) %>%
  summarise(medianArea = median(BsmtFinSF1), counts = n()) %>%
  arrange(medianArea) %>%
  ggplot(aes(x = reorder(BsmtFinType1, -medianArea), y = medianArea)) +
  geom_bar(stat = 'identity', fill='#00A4CC') +
  labs(x='BsmtFinType2', y='Median of BsmtFinSF2') +
  geom_text(aes(label = sort(medianArea)), vjust = -0.5) +
  theme_minimal() +
  scale_y_continuous(limits = c(0,850))

bsmt.fin.list = c('None' = 1, 'Unf' = 1, 'LwQ' = 2,'Rec'= 3, 'BLQ' = 4, 'ALQ' = 5, 'GLQ' = 6)
df.numeric = map.fcn(c('BsmtFinType1','BsmtFinType2'), bsmt.fin.list, df.numeric)
```

###Functional
```{r}
group.prices('Functional')
functional.levels = c('None'=0, 'Min2'=1, 'Maj2'=2, 'Min1'=3, 'Mod'=4, 'Maj1'=5, 'Sev'=6, 'Typ'=7)
df.numeric = map.fcn(c('Functional'), functional.levels, df.numeric)
```

###GarageFinish, Fences
```{r}
group.prices('GarageFinish')
garage.fin.list = c('None' = 0,'Unf' = 1, 'RFn' = 1, 'Fin' = 2)
df.numeric = map.fcn(c('GarageFinish'), functional.levels, df.numeric)

group.prices('Fence')
fence.levels = c('GdWo'=0, 'MnWw'=1, 'MnPrv'=2, 'None'=3, 'GdPrv'=4)
df.numeric = map.fcn(c('Fence'), fence.levels, df.numeric)

group.prices('MSSubClass')
MSSubClass.levels = c('30'=0, '180'=1, '90'=2, '190'=3, '50'=4, '85'=5,
                      '45'=6, '40'=7, '80'=8, '20'=9, '70'=10, '160'=11,
                      '75'=12, '120'=13, '60'=14)
df.numeric['NewerDwelling'] = as.numeric(MSSubClass.levels[as.character(df.combined$MSSubClass)])
```

```{r}
# need the SalePrice column
corr.df = cbind(df.numeric[1:1460,], train['SalePrice'])

# only using the first 1460 rows - training data
correlations = cor(corr.df)

corr.SalePrice = as.matrix(sort(correlations[,'SalePrice'], decreasing = TRUE))
corr.idx = names(which(apply(corr.SalePrice, 1, function(x) (x > 0.5 | x < -0.5))))

corrplot(as.matrix(correlations[corr.idx,corr.idx]), type = 'upper', method='color', addCoef.col = TRUE, tl.cex = 0.7,cl.cex = 0.7, number.cex=0.7)
```

```{r}
library(GGally)
ggpairs(corr.df[, corr.idx[1:6]])
```

